Arrows for Parallel Computations

### 2017-12-13 paper/content/arrows.tex
review of chapters 1-4 done

In this paper we will show that we can express parallel computations with Arrows and do not require Monads to do so. We also do not restrict the compatible Arrows to ones which have |ArrowApply| instances but only require the |ArrowChoice| and |ArrowLoop|. Because of this, we have a truly more general interface as compared to a monadic one.

Also note that, while we could have based our DSL on Profunctors as well, we chose Arrows. for this paper since they they are a more direct translation of pure functions than Profunctors. Note that Profunctors can easily be adapted to our interface as shown in Appendix~\ref{app:profunctorArrows}. However, they are a promising candidate for future improvement of our DSL.

% \caption{Add over arrows}
% \label{fig:addArrows}
% \end{figure}
% The benefit of using the |Arrow| typeclass is that any type which is shown to be an arrow can be used in conjunction with this newly created |add| combinator. Even though this example is quite simple, the power of the arrow interface immediately is clear: If a type is an arrow, it can automatically used together with every library that works on arrows. Compared to simple Monads, this enables us to write code that is more extensible, without touching the internals of the specific arrows.
% \\\\
% \textit{Note: In the definitions Hughes gave in his paper, the notation |a b c| for an arrow from |b| to |c| is used. We use the equivalent definition |arr a b| for an arrow from |a| to |b| instead, to make it easier to find the arrow type in type signatures.}
%

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:

>>>>

In this paper we will show that we can express parallel computations with Arrows and do not require Monads to do so. We also do not restrict the compatible Arrows to ones which have |ArrowApply| instances but instead only require instances for |ArrowChoice| and |ArrowLoop|. Because of this, we have a truly more general interface as compared to a monadic one.

Also note that, while we could have based our DSL on Profunctors as well, we chose Arrows. for this paper since they they are a more direct translation of pure functions than Profunctors. Note that Profunctors can easily be adapted to our interface as shown in Appendix~\ref{app:profunctorArrows}. However, they are a promising candidate for future improvements of our DSL.

### 2017-12-13 paper/content/basicSkeletons.tex
review of chapters 1-4 done

\subsubsection{Lazy |parEvalN|}
\begin{figure}[tb]
	\includegraphics[scale=0.7]{images/parEvalNLazy}
	\caption{Schematic depiction of |parEvalNLazy|.}
	\label{fig:parEvalNLazyImg}
\end{figure}
The function |parEvalN| is 100\% strict, which means that it fully evaluates all passed arrows. Sometimes this might not be feasible, as it will not work on infinite lists of functions like e.g. |map (arr . (+)) [1..]| or just because we need the arrows evaluated in chunks. |parEvalNLazy| (Figs.~\ref{fig:parEvalNLazyImg},~\ref{fig:parEvalNLazy}) fixes this. It works by first chunking the input from |[a]| to |[[a]]| with the given |ChunkSize| in |arr (chunksOf chunkSize)|. These chunks are then fed into a list |[arr [a] [b]]| of parallel arrows created by feeding chunks of the passed |ChunkSize| into the regular parEvalN by using |listApp|. The resulting |[[b]]| is lastly converted into |[b]| with |arr concat|.
\begin{figure}[t]
\begin{code}
parEvalNLazy :: (ArrowParallel arr a b conf, ArrowChoice arr, ArrowApply arr) =>
	conf -> ChunkSize -> [arr a b] -> (arr [a] [b])
parEvalNLazy conf chunkSize fs =
	arr (chunksOf chunkSize) >>>
	listApp fchunks >>>
	arr concat
	where fchunks = map (parEvalN conf) $ chunksOf chunkSize fs
\end{code} %$ %% formatting
\caption{Definition of |parEvalNLazy|.}
\label{fig:parEvalNLazy}
\end{figure}

\subsubsection{Heterogeneous tasks}
\begin{figure}[tb]
	\includegraphics[scale=0.7]{images/parEval2}
	\caption{Schematic depiction of |parEval2|.}
	\label{fig:parEval2Img}
\end{figure}
We have only talked about the parallelization arrows of the same type until now. But sometimes we want to parallelize heterogeneous types as well. However, we can implement such a |parEval2| combinator (Figs.~\ref{fig:parEval2Img},~\ref{fig:parEval2}) which combines two arrows |arr a b| and |arr c d| into a new parallel arrow |arr (a, c) (b, d)| quite easily with the help of the |ArrowChoice| type class. The idea is to use the |+++| combinator which combines two arrows |arr a b| and |arr c d| and transforms them into |arr (Either a c) (Either b d)| to get a common arrow type that we can then feed into |parEvalN|.

>>>>

\subsubsection{Lazy |parEvalN|}
\begin{figure}[tb]
	\includegraphics[scale=0.7]{images/parEvalNLazy}
	\caption{Schematic depiction of |parEvalNLazy|.}
	\label{fig:parEvalNLazyImg}
\end{figure}
The function |parEvalN| is 100\% strict, which means that it fully evaluates all passed arrows. Sometimes this might not be feasible, as it will not work on infinite lists of functions like e.g. |map (arr . (+)) [1..]| or just because we need the arrows evaluated in chunks. |parEvalNLazy| (Figs.~\ref{fig:parEvalNLazyImg},~\ref{fig:parEvalNLazy}) fixes this. It works by first chunking the input from |[a]| to |[[a]]| with the given |ChunkSize| in |arr (chunksOf chunkSize)|. These chunks are then fed into a list |[arr [a] [b]]| of parallel arrows created by feeding chunks of the passed |ChunkSize| into the regular parEvalN by using |listApp|. The resulting |[[b]]| is lastly converted into |[b]| with |arr concat|.
\begin{figure}[t]
\begin{code}
parEvalNLazy :: (ArrowParallel arr a b conf, ArrowChoice arr, ArrowApply arr) =>
	conf -> ChunkSize -> [arr a b] -> (arr [a] [b])
parEvalNLazy conf chunkSize fs =
	arr (chunksOf chunkSize) >>>
    evalN fchunks >>>
    arr concat
    where
      fchunks = map (parEvalN conf) (chunksOf chunkSize fs)
\end{code} %$ %% formatting
\caption{Definition of |parEvalNLazy|.}
\label{fig:parEvalNLazy}
\end{figure}

\subsubsection{Heterogeneous tasks}
\begin{figure}[tb]
	\includegraphics[scale=0.7]{images/parEval2}
	\caption{Schematic depiction of |parEval2|.}
	\label{fig:parEval2Img}
\end{figure}
We have only talked about the parallelisation of arrows of the same type until now. But sometimes we want to parallelize heterogeneous types as well. We can implement such a |parEval2| combinator (Figs.~\ref{fig:parEval2Img},~\ref{fig:parEval2}) which combines two arrows |arr a b| and |arr c d| into a new parallel arrow |arr (a, c) (b, d)| quite easily with the help of the |ArrowChoice| type class. The idea is to use the |+++| combinator which combines two arrows |arr a b| and |arr c d| and transforms them into |arr (Either a c) (Either b d)| to get a common arrow type that we can then feed into |parEvalN|.

### 2017-12-13 paper/content/utilityFunctions.tex
review of chapters 1-4 done

Following is a theoretical way of using Profunctors together with our DSL.
\footnote{for additional information on the typeclasses, see: \url{https://hackage.haskell.org/package/profunctors-5.2.1/docs/Data-Profunctor.html}}

\begin{code}
instance (Category p, Strong p) => Arrow p where
  arr f = dimap id f id
  first = first'

@@

instance (Category p, Strong p, Choice p) => ArrowChoice p where
  left = left'
\end{code}

>>>>

In Fig.~\ref{fig:profunctorArrow} we show how specific Profunctors can be turned into Arrows.

\begin{figure}[h]
\begin{code}
instance (Category p, Strong p) => Arrow p where
  arr f = dimap id f id
  first = first'

@@

instance (Category p, Strong p, Choice p) => ArrowChoice p where
  left = left'
\end{code}
\caption{Profunctors as Arrows}
\footnote{for additional information on the typeclasses see: \url{https://hackage.haskell.org/package/profunctors-5.2.1/docs/Data-Profunctor.html}}}
\label{fig:profunctorArrow}
\end{figure}
